Gait Segmentation Model for Rotation Time Series Collected by a Wearable Sensor using Machine Learning

### 2024-12-12 article_segmentation.qmd
Properly indent blocks.

Studies show good results when using machine learning to classify gait of patients having trouble walking, for instance patients with Parkinson Disease [@tahir2012parkinson;@wahid2015classification], or to extract gait parameters [@rampp2014inertial;@hannink2016sensor]. These studies mostly use Artificial Neural Networks but do not address the problem of the segmentation of the signal recorded by the sensors. Similarly, in @farah2019design, a logistic regression tree is used to classify gait phases in a stride but not to directly segment the signal into strides.

For the segmentation matter, Recurrent Neural Networks have shown good results for identifying Heel-Strike and Toe-off events with pressure sensors, accelerations and Euler angles [@prado2019gait]. Another study compared unsupervised machine learning (k-means) with supervised one (SVM and CNN), showing that CNNs were performing the best to predict stance and swing phases [@potluri2019machine]. Lastly, in another study, 24 time series from three sensors’ accelerometers and gyroscopes were used as input into a 6 layers CNN to estimate the likelihood of the corresponding input sample, given a specific gait event [@gadaleta2019deep]. For instance, the signal can be segmented with the initial contact of the right foot if this gait event is chosen in the model. This method was shown to perform better than a Wavelet Transform algorithm.

>>>>

: Studies show good results when using machine learning to classify gait of patients having trouble walking, for instance patients with Parkinson Disease [@tahir2012parkinson;@wahid2015classification], or to extract gait parameters [@rampp2014inertial;@hannink2016sensor]. These studies mostly use Artificial Neural Networks but do not address the problem of the segmentation of the signal recorded by the sensors. Similarly, in @farah2019design, a logistic regression tree is used to classify gait phases in a stride but not to directly segment the signal into strides.

: For the segmentation matter, Recurrent Neural Networks have shown good results for identifying Heel-Strike and Toe-off events with pressure sensors, accelerations and Euler angles [@prado2019gait]. Another study compared unsupervised machine learning (k-means) with supervised one (SVM and CNN), showing that CNNs were performing the best to predict stance and swing phases [@potluri2019machine]. Lastly, in another study, 24 time series from three sensors’ accelerometers and gyroscopes were used as input into a 6 layers CNN to estimate the likelihood of the corresponding input sample, given a specific gait event [@gadaleta2019deep]. For instance, the signal can be segmented with the initial contact of the right foot if this gait event is chosen in the model. This method was shown to perform better than a Wavelet Transform algorithm.
